---
title: "10_year_risk_of_future_coronary_disease"
author: "Qing Wen"
date: "2020/12/2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(skimr)
# read in the original dataset
heart_data_original <- read_csv("C:/Users/10128/Desktop/STA304/Final project/framingham.csv")
skim(heart_data_original)

# take out two variables and remove the missing values in the rest of the observations
heart_data <- heart_data_original %>% 
  select(-c(education, glucose)) %>% 
  na.omit() 

skim(heart_data)

# save the cleaned dataset 
write_csv(heart_data, "C:/Users/10128/Desktop/STA304/Final project/cleaned_data.csv")
```

Despite of some missing values, the original data is pretty clean and easy to interpret. 

With the objective of our study, we decide to remove the variable **education** as we do not think it has an as strong relationship with predicting the 10-yaer-risk of having heart disease. Also, variable **glucose** is removed from the dataset as it contains too many missing values(388 missing observations). 

After taking out these two variables, we then omit all the missing values in the remaining observations, which left us with 4088 observations in total.
